---
output: github_document
editor_options:
  markdown:
    wrap: sentence
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
# Turn on ANSI colors
asciicast::init_knitr_engine(
  startup = quote({
    library(cli)
    set.seed(1) }),
  echo = FALSE,
  echo_input = FALSE)
```
 
# `cTOST` Overview

This repository holds the `cTOST` R package. This package contains the function `tost` which provides an assessment of equivalence in the univariate framework based on the state-of-the-art Two One-Sided Tests (TOST). In addition, the package contains the functions `atost` and `dtost`, two corrective procedures applied to the TOST in the univariate framework in order to ensure the preservation of the Type I error rate at the desired nominal level and a uniform increase in power. These two functions output an assessment of equivalence in the univariate framework after their respective corrections is applied. More details can be found in @boulaguiem2023finite that you can access via this [link](https://www.biorxiv.org/content/10.1101/2023.03.11.532179v3).

# Install Instructions

The `cTOST` package is only available on GitHub at the moment. It is subject to ongoing updates that may lead to stability issues.  

In order to install the package, it is required to pre-install the `devtools` dependency. Run the following command if you do not have it already installed:
```{r, eval=FALSE}
install.packages("devtools")
```
The package is then installed with the following command: 
```{r, eval=FALSE}
devtools::install_github("cTOST")
```

# Equivalence Assessment 

## Skin dataset

In order the demonstrate the use of the functions described above, we will use it on a dataset included in the package. The dataset was obtained using the cutaneous bioequivalence method detailed in @quartier2019cutaneous and was provided by the same authors. It contains $17$ pairs of comparable porcine skin samples on which measurements of econazole nitrate deposition (an antifungal medication used to treat skin infections) were collected using two (supposedly) bioequivalent products. These measurements were then considered on the logarithmic scale and saved as an RData file called `skin`. 

After installing the package, you can simply print the dataset by running `skin` on your console (the dataset has been lazy loaded into the package). 

```{r, eval=TRUE}
library(cTOST)
# data(skin) # not necessary to run this line as the dataset is lazy loaded in the package
head(skin)
```

The following code allows to reproduce Figure 1 of @boulaguiem2023finite:

```{r, eval=TRUE}
par(mar=c(2.5,5,0,0))

# Empty plot
plot(1,1,pch="",axes=FALSE,xlab="",ylab="",main="",
     ylim=c(min(unlist(skin)),log(4000)),xlim=c(0.75, 2.25))

# Plotting and linking the paired data points of each group
for(i in 1:nrow(skin)){
  points(c(1,2),skin[i,],col=paste0(gray(0.1),50),pch=16)
  segments(1,skin[i,1],2,skin[i,2],
           col=paste0(gray(0.1),50))
}
axis(1, at = c(1,2), c("Reference", "Generic"),tick=FALSE)
axis(2, at = log(c(250,500,1000,2000,4000)), c(250,500,1000,2000,4000), las=2)
axis(2, at = mean(c(min(unlist(skin)),log(4000))), expression(paste("ECZ deposition (ng/cm"^2*")")),padj=-4.5, tick = FALSE)
```

Let us now extract from the dataset the components needed to run the functions `atost` and `dtost` We would need the difference of means between the two groups, the number of degrees of freedom, and the standard error:

```{r, eval=TRUE}
# Difference in means between the two groups
theta_hat <- diff(apply(skin,2,mean)) 

# Number of degrees of freedom
nu <- nrow(skin)-1 

# Standard error
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
```

Considering a significance level of $5\%$ and an equivalence limit at $\text{delta}=log(1.25)$, the `tost`, `atost` and `dtost` functions are used as follows:

## Standard TOST

```{r, eval = F}
res_tost = tost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_tost
```

```{asciicast}
library(cTOST)
data(skin)
theta_hat = diff(apply(skin,2,mean))
nu = nrow(skin) - 1
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
res_tost = tost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_tost
```

## $\alpha-TOST$

```{r, eval = F}
res_atost = atost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_atost
```

```{asciicast}
library(cTOST)
data(skin)
theta_hat = diff(apply(skin,2,mean))
nu = nrow(skin) - 1
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
res_atost = atost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_atost
```

```{r, eval=FALSE}
compare_to_tost(res_atost)
```

```{asciicast}
library(cTOST)
data(skin)
theta_hat = diff(apply(skin,2,mean))
nu = nrow(skin) - 1
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
res_atost = atost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
compare_to_tost(res_atost)
```

## $\delta-TOST$

```{r, eval = F}
res_dtost = dtost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_dtost
```

```{asciicast}
library(cTOST)
data(skin)
theta_hat = diff(apply(skin,2,mean))
nu = nrow(skin) - 1
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
res_dtost = dtost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
res_dtost
```

```{r, eval=FALSE}
compare_to_tost(res_dtost)
```

```{asciicast}
library(cTOST)
data(skin)
theta_hat = diff(apply(skin,2,mean))
nu = nrow(skin) - 1
sig_hat = sd(apply(skin,1,diff))/sqrt(nu)
res_dtost = dtost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
compare_to_tost(res_dtost)
```

## Interval Inclusion Principal with the $\alpha$-TOST

To visually assess equivalence with the interval inclusion principal, we reproduce Figure 2 of @boulaguiem2023finite with the following code: 

```{r include=TRUE}
res_atost = atost(theta = theta_hat, sigma = sig_hat, nu = nu,
              alpha = 0.05, delta = log(1.25))
```

```{r, eval=TRUE}
# Nominal level and equivalence bounds
alpha = 0.05
delta = log(1.25)

# Empty plot
plot(NA, axes = F, xlim = c(-0.25, 0.25), ylim = c(0.5, 2.5),
     xlab = " ", ylab = " ")

# Corrected level used by aTOST
alpha_star = res_atost$corrected_alpha

# TOST's CI upper and lower bound
a1 = theta_hat + sig_hat*qt(alpha, df = nu)
a2 = theta_hat - sig_hat*qt(alpha, df = nu)

# aTOST's CI upper and lower bound 
b1 = theta_hat + sig_hat*qt(alpha_star, df = nu)
b2 = theta_hat - sig_hat*qt(alpha_star, df = nu)

# Set the colors and their transparent counterpart
cols=c("coral","palegreen2")
t.cols=sapply(c("coral","palegreen2"),function(x){
  rgb.val <- col2rgb(x)
  rgb(rgb.val[1], rgb.val[2], rgb.val[3],
      max = 255,
      alpha = 127.5)
})

# Plotting the TOST and aTOST CIs
d = 0.1
polygon(c(a1, a2, a2, a1), c(1-d, 1-d, 1+d, 1+d) + 1, border = NA,
        col = t.cols[1])
polygon(c(delta, a2, a2, delta), c(1-d, 1-d, 1+d, 1+d) + 1, border = NA,
        col = cols[1], density = 20, angle = 45)
polygon(c(delta, a2, a2, delta), c(1-d, 1-d, 1+d, 1+d) + 1, border = NA,
        col = cols[1], density = 20, angle = -45)
polygon(c(b1, b2, b2, b1), c(1-d, 1-d, 1+d, 1+d), border = NA,
        col = t.cols[2])

# Plotting the difference in means
points(theta_hat, 2, col = cols[1], pch = 16, cex = 2.5)
points(theta_hat, 1, col = cols[2], pch = 16, cex = 2.5)

# Adding text for readability
text(0.15, 2 + 1.75*d, "TOST CI")
text(0.15, 1 + 1.75*d, expression(paste(alpha,"TOST CI")))

text(0.22, 2 - 1.75*d, expression(paste("outside (-",delta,", ",delta,")")), cex = 0.8)
abline(v = c(-0.3, -log(1.25), -0.1, 0, 0.1, log(1.25), 0.3),
       col = "grey65", lty = 2)
axis(1, at = c(-0.3, -log(1.25), -0.1, 0, 0.1, log(1.25), 0.3),
     labels = c("-0.3", expression(paste("-",delta)), "-0.1", "0", "0.1",
                expression(delta), "0.3"),padj=-0.15,cex.axis=1.3)
mtext(expression(theta), side = 1, line = 2.35, cex = 1.4)

``` 

# References









