---
output: github_document
editor_options:
  markdown:
    wrap: sentence
bibliography: references.bib
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit this file -->

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
# Turn on ANSI colors
asciicast::init_knitr_engine(
  startup = quote({
    library(cli)
    set.seed(1) }),
  echo = FALSE,
  echo_input = FALSE)
```
 
<!-- badges: start -->
[![R-CMD-check](https://github.com/yboulag/cTOST/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yboulag/cTOST/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# `cTOST` Overview

This repository holds the `cTOST` R package. This package contains the function `tost` which provides an assessment of equivalence in the univariate framework based on the state-of-the-art Two One-Sided Tests (TOST). In addition, the package contains the functions `atost` and `dtost`, two corrective procedures applied to the TOST in the univariate framework in order to ensure the preservation of the Type I error rate at the desired nominal level and a uniform increase in power. These two functions output an assessment of equivalence in the univariate framework after their respective corrections is applied. More details can be found in @boulaguiem2023finite that you can access via this [link](https://www.biorxiv.org/content/10.1101/2023.03.11.532179v3).

# Install Instructions

The `cTOST` package is available on GitHub at the moment. It is subject to ongoing updates that may lead to stability issues.  

In order to install the package, it is required to pre-install the `devtools` dependency. Run the following command if you do not have it already installed:
```{r, eval=FALSE}
install.packages("devtools")
```
The package is then installed with the following command: 
```{r, eval=FALSE}
devtools::install_github("yboulag/cTOST")
```
Note that Windows users are assumed that have Rtools installed (if this is not the case, please visit this [link](https://cran.r-project.org/bin/windows/Rtools/).

# How to cite

```{r, eval=FALSE}
@Manual{boulaguiem2023ctost,
  title = {cTOST: Finite Sample Correction of The TOST in The Univariate Framework},
  author = {Younes Boulaguiem and Stéphane Guerrier and Dominique-Laurent Couturier},
  year = {2023},
  note = {R package version 1.0.0},
  url = {https://github.com/yboulag/cTOST},
}
```

# License

The license this source code is released under is the GNU AFFERO GENERAL PUBLIC LICENSE (AGPL) v3.0. Please see the LICENSE file for full text. Otherwise, please consult [GNU](https://www.gnu.org/licenses/agpl-3.0.en.html) which will provide a synopsis of the restrictions placed upon the code.

# References









